---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dtreg

<!-- badges: start -->
<!-- badges: end -->

The goal of dtreg is to help the user interact with various data type registries (DTRs).
The user can load a DTR schema as an R object and create their own instance of the schema.
This instance can be written as a machine-actionable JSON-LD file. 

## Installation

You can install the development version of dtreg from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("OlgaLezhnina/dtreg")
```

## Example

This example shows you how to work with a schema from a DTR; you need to know the DOI of the schema.
The schema we use as an example belongs to the ePIC DTR with DOI "https://doi.org/21.T11969/74bc7748b8cd520908bc".

```{r example}
library(dtreg)
## load the datatype schema you need
dt <- dtreg::load_datatype("https://doi.org/21.T11969/74bc7748b8cd520908bc")
## look at schemata to select the one(s) you intend to use
names(dt)
## check available fields for your schema
dtreg::show_fields(dt$inferential_test_output())
## create your own instance by filling the fields of your choice
my_inst <- dt$inferential_test_output(label = "my_results")
## write the instance in JSON-LD format
my_json <- dtreg::to_jsonld(my_inst)
```

For more information, please see XXX.
