---
title: "Introduction to dtreg"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to dtreg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dtreg)
```
You can use dtreg for the following tasks:

* Interact with data type registries (DTRs): currently, the ePIC and the ORKG DTRs.

* Write down the results of your data analysis in accordance with a DTR schema.

* Convert the results into a machine-readable JSON-LD file. 

## Load a DTR schema 

To load a DTR schema, you need to know it identifier.
For the EPIC, it is the datatype DOI, such as "https://doi.org/21.T11969/ff5e3f857788d20dd1aa".
For the ORKG, it is the template URL, such as "https://incubating.orkg.org/template/R836000".
If a valid identifier is used, you get an R object containing information about the DTR schema.

```{r}
dt <- dtreg::load_datatype("https://doi.org/21.T11969/ff5e3f857788d20dd1aa")
```
In addition to the schema you requested, you get related schemata, which could be  useful for reporting your results.
You can look at the list of these schemata:

```{r}
names(dt)
```
## Create an instance 

To write down the data analysis results, dtreg uses R6 classes. Therefore, you need to create an instance of a specific class.

### Fields 

For doing that, you first need to know which fields you can use: 

```{r}
dtreg::show_fields(dt$statistical_test_of_difference())
```

For example, there is the field "label". If your instance included only a label, it would be:


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

```{r}
labelled_inst <- dt$statistical_test_of_difference(label = "my_test_results")
```

However, the results of your data analysis are more complex than this.
Thus, it is recommendable to consult the help page XXX which explains whether a field requires a schema or a specific type of data (e.g., numeric).  

### String, numeric, and data frame

Most frequently used types are string, numeric, and data frame.
Strings are used for labels and comments, and URLs are also presented as strings:

```{r}
method_URL <- dt$software_method(
                 has_support_URL = "https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/wilcox.test")
```

Numeric values are also frequently used:

```{r}
sample_size <- dt$sample_size(has_specified_numeric_value = 148)
```

Some fields, such as "has_format", require a data frame: 

```{r}
my_dataframe <- data.frame(p = 0.026, t = 2.3, df = 48, stringsAsFactors = FALSE)
output_dataframe <- dt$inferential_test_output(has_format = my_dataframe)
```

By default, a data frame is assigned a generic name "Table". If you want to give your data frame a name, you can use a tuple. Please always include first the data frame and then the name as a string in the tuple. 

```{r}
my_tuple <- sets::tuple(my_dataframe, "the Wilcoxon results")
output_tuple <- dt$inferential_test_output(has_format = my_tuple)
```

There was some code I removed because of the whatever problem with the sets. 

### More than one input in a field

Sometimes a few objects should be written in one field.
In this case simply concatenate them:

```{r}
var_1 <- dt$statistical_variable(label = "var_1")
var_2 <- dt$statistical_variable(label = "var_2")
two_vars <- dt$statistical_test_of_difference(evaluates = c(var_1, var_2))
```

### Nested structure

In the example above you can see that a field of a schema might require another schema. It can be also written this way, depending on what is more convenient for you: 


```{r}
two_vars <- 
  dt$statistical_test_of_difference(
    evaluates = c(
      dt$statistical_variable(label = "var_1"),
      dt$statistical_variable(label = "var_2")
    )
  )
```

This nested structure is informed by machine-readability of the data. The further versions of dtreg will suggest very simple wrappers to write the results for you.

### Gather all info to write an instance 

Let us write down all components of the data analysis results and gather them into an instance. 

## Convert the instance into a JSON-LD file

This stage is simple:


```{r}
my_json <- dtreg::to_jsonld(labelled_inst)
```
